Estimating the Initial Susceptible Fraction in Dengue Dynamics

### 2015-06-23 draft_lm1.tex
minor comment 6 closes #34

Quantifying the attack ratio of disease is key to epidemiological inference and 
Public Health planning.
For multi-serotype pathogens, however, different levels of 
serotype-specific immunity make it difficult to assess the population at risk. 
In this paper we propose a Bayesian method for estimation of the attack ratio 
of an epidemic and the initial fraction of susceptibles using aggregated 
incidence data. 
We derive the probability distribution of the effective reproductive number, 
\rr, and use MCMC to obtain posterior 
distributions of the parameters of a single-strain SIR transmission model with 
time-varying force of infection.
Our method is showcased in a data set consisting of 18 years of dengue 
incidence in the city of Rio de Janeiro, Brazil.
We demonstrate that it is possible to learn about the initial fraction of 
susceptibles and the attack ratio even in the absence of serotype specific data.
On the other hand, the information provided by this approach is limited, 
stressing the need for detailed serological surveys to characterise the 
distribution of serotype-specific immunity in the population.

@@

We take a Bayesian approach to the estimation of $S_0$.
First the incidence time series was divided into $J=13$ 
epidemic windows that corresponded to significant raises in incidence and 
normalized to lie on the $[0,1]$ interval.
For a given interval $j = \{ t_j^{\text{start}}, t_j^{\text{end}} \} $ we 
observe an incidence time series $\mathbf{Y_{j}}$.
We are thus interested in the posterior distribution
\begin{equation}
 \label{eq:S0post}
 p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, \tau 
)\pi(S_{0j}) 
\end{equation}
The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is assumed to be a Normal distribution 
with fixed variance $\sigma^2$.
In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
obtained as described above and fixed $\tau = 1/7\: \text{days}^{-1}$.
To complete the inference, we need to specify prior distributions for the 
parameters of interest.
We place a flat $\text{Beta}(1, 1)$ prior on $S_{0j}\:\forall j$

>>>>

Quantifying the attack ratio of disease is key to epidemiological inference and 
public health planning.
For multi-serotype pathogens, however, different levels of 
serotype-specific immunity make it difficult to assess the population at risk. 
In this paper we propose a Bayesian method for estimation of the attack ratio 
of an epidemic and the initial fraction of susceptibles using aggregated 
incidence data. 
We derive the probability distribution of the effective reproductive number, 
\rr, and use MCMC to obtain posterior 
distributions of the parameters of a single-strain SIR transmission model with 
time-varying force of infection.
Our method is showcased in a data set consisting of 18 years of dengue 
incidence in the city of Rio de Janeiro, Brazil.
We demonstrate that it is possible to learn about the initial fraction of 
susceptibles and the attack ratio even in the absence of serotype specific data.
On the other hand, the information provided by this approach is limited, 
stressing the need for detailed serological surveys to characterise the 
distribution of serotype-specific immunity in the population.

@@

We take a Bayesian approach to the estimation of $S_0$.
First the incidence time series was divided into $J=13$ 
epidemic windows that corresponded to significant raises in incidence and 
normalized to lie on the $[0,1]$ interval.
For a given interval $j = \{ t_j^{\text{start}}, t_j^{\text{end}} \} $ we 
observe an incidence time series $\mathbf{Y_{j}}$.
We are thus interested in the posterior distribution
\begin{equation}
 \label{eq:S0post}
 p(S_{0j}|\mathbf{Y_{j}}) \propto l(\mathbf{Y_{j}}|S_{0j}, R_t, \tau 
)\pi(S_{0j}) 
\end{equation}
The likelihood $l(\mathbf{Y_{j}}|\cdot)$ is a approximated by a Normal 
distribution with fixed variance $\sigma^2$. This approximation is a numerical 
convenience since it confers better stability to the MCMC sampling.
In this estimation procedure we kept $R_t$ fixed at fixed at the posterior mean 
obtained as described above and fixed $\tau = 1/7\: \text{days}^{-1}$.
To complete the inference, we need to specify prior distributions for the 
parameters of interest.
We place a flat $\text{Beta}(1, 1)$ prior on $S_{0j}\:\forall j$
